---
title: "milestone4"
author: "Ryan Graff"
date: "October 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lintr)
library(gt)
library(readr)
library(dplyr)
library(glue)
library(knitr)
library(styler)
library(tidyr)
library(readxl)
library(purrr)
library(tibble)
library(stringr)
library(janitor)
library(ggplot2)
library(tidyverse)
```


```{r data import, include=FALSE, echo=FALSE}
draft <- read_xlsx("data/draft_data.xlsx", skip = 1) %>%
  clean_names()
# Reading in my data, and cleaning up the names so they will be easier to manipulate
```

```{r plot_data, echo=FALSE}
# gathering the data I want to use for my plot
plot <- draft %>% 
  #I want the first pick of the 1st and 2nd rounds, so I filter accordingly
  filter(rd == 1 | 2, pk == 1) %>% 
  # Now I grab the year, round, VORP, and PPG of the draft picks, and am now ready to plot
  select(year, rd, vorp, pts_18)
  
```

# Plot
```{r graphic, echo=FALSE, warning=FALSE}
# Creating the plot, I chose a scatter plot because I wanted each pick to have its own data point.  I added in ppg for size of the dot in order to convey more information about the ability of each pick.
ggplot(plot, aes(x = year, y = vorp, size = pts_18)) + geom_point(aes(color = factor(rd), size = pts_18)) + labs(title = "The Title of 'First Pick' is Not Created Equal", subtitle = "VORPs and PPG of the 1st picks in the NBA draft's 1st and 2nd rounds, 1987-2018") + xlab("Year") + ylab("VORP") + labs(color = "Round", size = "Points Per Game (PPG)")

```

# Description: 

  VORP is an advanced basketball metric that shows how much better (or worse) the player in question is compared to the "league replacement", also known as a theoretical player making minimum salary or one that has just been brought from the minor leagues to the pros. This graph illustrates that while the individual VORPs of the 1st round, 1st overall picks are all over the place, they are generally at or above 20, and almost always better than the first pick of the second round of their respective draft.  Meanwhile, the first picks of the second round don't have VORPs much better than that of a replacement player, which is -2. 
  The size of the dots correspond to the player's average points per game.  The first rounders continually average more points per game than the second rounders.  One can also see that higher points per game usually corresponds with a higher VORP.
  
  